### 2017-11-28 paper/sections/1_introduction.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot spots
that offer significant opportunity for optimization. We use the data management
language and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different application and storage system.  Our exploration of this space results in a
dynamically sized cache policy that sacrifices negligible performance while
using 32-66\% less memory than the default ParSplice configuration.

@@

Our example application, ParSplice~\cite{perez:jctc20150parsplice}, is
representative of an important class applications with similar working set
behaviors that extensively use software-based caches.  ParSplice is a molecular
dynamics simulation that uses a hierarchy of caches and a single persistent
key-value store to store both observed minima across a molecule's equation of
motion (EOM) and the hundreds or thousands of partial trajectories calculated
each second during a parallel job.  This workload is pervasive across
simulations that rely on a mesh-based decomposition of a physical region and
result in millions or billions of mesh cells, where each cell contains
materials, pressures, temperatures and other characteristics that are required
to accurately simulate phenomena of interest.  The fine-grained data annotation
capabilities provided by key-value storage is a natural match for these types
of scientific simulations.  Unfortunately, simulations of this size saturate
the capacity and bandwidth capabilities of a single node so we need more
effective data management techniques.

@@

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, we find this threshold
works well for HPC applications.  For reference, a 1GB cache for a distributed
file system is too large in LANL deployments.}. Furthermore, ParSplice deploys
a cache per 300 worker processes, so large simulations need more caches and
will use even more memory.  Users can configure ParSplice to evict data when
the cache reaches a threshold but this solution requires tuning and parameter
sweeps; the ``Cache (too small)" curve in Figure~\ref{fig:cache-management}
shows how a poorly configured cache can save memory but at the cost of
performance, which is shown by the text annotation to the right.  Even worse,
this threshold changes with different initial configurations and cluster setups
so tuning needs to be done for all system permutations.  Our dynamically sized
cache, shown by the thick line in Figure~\ref{fig:cache-management}, detects
key access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache with a
negligible performance degradation (within the baseline's standard deviation),
works for a variety of initial conditions, and could generalize to similar
applications.

@@

This last contribution is explored in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}, where we try a range of policies from different
disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
that the collection of policies we design for cache management in ParSplice
are very similar to the policies used to load balance metadata in the Ceph file
system (CephFS) suggesting that there is potential for automatically adapting
and generating policies dynamically.

>>>>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot spots
that offer significant opportunity for optimization. We use the data management
language and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different application and storage system.  Our exploration of this space results in a
dynamically sized cache policy that does not sacrifice any performance while
using 32-66\% less memory than the default ParSplice configuration.

@@

Our example application, ParSplice~\cite{perez:jctc20150parsplice}, is
representative of an important class of applications with similar working set
behaviors that extensively use software-based caches.  ParSplice is a molecular
dynamics simulation that uses a hierarchy of caches and a single persistent
key-value store to store both observed minima across a molecule's equation of
motion (EOM) and the hundreds or thousands of partial trajectories calculated
each second during a parallel job.  This workload is pervasive across
simulations that rely on a mesh-based decomposition of a physical region and
result in millions or billions of mesh cells, where each cell contains
materials, pressures, temperatures and other characteristics that are required
to accurately simulate phenomena of interest.  The fine-grained data annotation
capabilities provided by key-value storage is a natural match for these types
of scientific simulations.  Unfortunately, simulations of this size saturate
the capacity and bandwidth capabilities of a single node so we need more
effective data management techniques.

@@

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, this threshold works well
for HPC applications.  For reference, a 1GB cache for a distributed file system
is too large in LANL deployments.}. Furthermore, ParSplice deploys a cache per
300 worker processes, so large simulations need more caches and will use even
more memory.  Users can configure ParSplice to evict data when the cache
reaches a threshold but this solution requires tuning and parameter sweeps; the
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the cost of performance, which
is shown by the text annotation to the right.  Even worse, this threshold
changes with different initial configurations and cluster setups so tuning
needs to be done for all system permutations.  Our dynamically sized cache,
shown by the thick line in Figure~\ref{fig:cache-management}, detects key
access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache
without any performance degradation, works for a variety of initial conditions,
and could generalize to similar applications.

@@

This last contribution is explored in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}, where we try a range of policies from different
disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
that the collection of policies we design for cache management in ParSplice are
very similar to the policies used to load balance metadata in the Ceph file
system (CephFS~\cite{weil:osdi2006-ceph}) suggesting that there is potential
for automatically adapting and generating policies dynamically.

### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsubsection*{Experimental Setup} All experiments ran on Trinitite, a Cray
XC40 with 32 Intel Haswell 2.3GHz cores per node.  Each node has 128GB of RAM
and our goal is to limit the size of the cache to 3\% of RAM . Note that this
is an addition to the 30GB that ParSplice uses to manage other ranks on the
same node.  A single Cray node produced trajectories that are \(5\times\) times
longer than our 10 node CloudLab clusters and \(25\times\) longer than our 10
node cluster at UC Santa Cruz. As a result, it reaches different job phases faster and
gives us a more comprehensive view of the workload. The performance gains
compared to the commodity clusters have more to do with memory/PCI bandwidth
than network.

The scalability experiment uses 1 splicer, 1 persistent database, 1 cache
process, and up to 2 workers. We scale up to 1024 tasks, which spans 32 nodes
and disable hyper-threading because we experience unacceptable variability in
performance. For the rest of the experiments, we use 8 nodes, 1 splicer, 1
persistent database, 1 cache process, 1 worker, and up to 256 tasks.  The
keyspace analysis that follows is for the cache on the worker node, which is
circled in Figure~\ref{fig:parsplice}.

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
  \caption{The keyspace size is small but must satisfy many reads as workers
  calculate new segments. It is likely that we will need
  more than one node to manage segment coordinates when we scale the system or jobs up.
  \label{fig:methodology-keyspace}}
\end{figure}

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
  \caption{The keyspace activity for ParSplice using two different growth
rates.  The line shows the rate that EOM minima values are retrieved from the
key-value store (\(y1\) axis) and the points along the bottom show the number
of keys accessed in a 1 second sliding window (\(y2\) axis).
\label{fig:motivation-regimes}}
\end{figure}

@@

\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
keyspace size (text annotations) and request load (bars) after a one hour run
with a different number of tasks(\(x\) axis). While the keyspace size and
capacity is relatively modest the memory usage scales linearly with the number
of tasks. This is a problem if we want to scale to Trinitite's 3000 cores.
Furthermore, the size of the keyspace also increases linearly with the length
of the run.  Extrapolating these results puts an 8 hour run across all 100
Trinitite nodes at 8GB for one cache.  This memory utilization easily eclipses
the 3\% memory usage per node threshold we set earlier, even without factoring
in the usage from other workers.

@@

The bars in Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many
reads (\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same
key for extended periods because the trajectory can remain stuck in so-called
superbasins composed of tightly connected sets of states. In this case, many
trajectory segments with the same coordinates are needed before the trajectory
moves on.  Writes only occur for the final state of segments generated by
tasks; their magnitude is smaller than reads because the caches ignore
redundant write requests. The number of read and write requests are highest at
the beginning of the run when tasks generate segments for the same state, which
is computationally cheap (this motivates Section~\S\ref{sec:arch-specific}).

>>>>

\subsubsection*{Experimental Setup} All experiments ran on Trinitite, a Cray
XC40 with 32 Intel Haswell 2.3GHz cores per node.  Each node has 128GB of RAM
and our goal is to limit the size of the cache to 3\% of RAM . Note that this
is an addition to the 30GB that ParSplice uses to manage other ranks on the
same node.  The scalability experiment uses 1 splicer, 1 persistent database, 1
cache process, and up to 2 workers. We scale up to 1024 tasks, which spans 32
nodes and disable hyper-threading because we experience unacceptable
variability in performance. For the rest of the experiments, we use 8 nodes, 1
splicer, 1 persistent database, 1 cache process, 1 worker, and up to 256 tasks.
The keyspace analysis that follows is for the cache on the worker node, which
is circled in Figure~\ref{fig:parsplice}.

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
  \caption{The keyspace size is small but must satisfy many reads as workers
  calculate new segments. Memory usage scales linearly, so It is likely that we will need
  more than one node to manage segment coordinates when we scale the system or jobs up.
  \label{fig:methodology-keyspace}}
\end{figure}

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
  \caption{The key activity for ParSplice using two different growth
rates.  The line shows the rate that EOM minima values are retrieved from the
key-value store (\(y1\) axis) and the points along the bottom show the number
of keys accessed in a 1 second sliding window (\(y2\) axis).\label{fig:motivation-regimes}}
\end{figure}

@@

\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
keyspace size (text annotations) and request load (bars) after a one hour run
with a different number of tasks (\(x\) axis). While the keyspace size and
capacity is modest the memory usage scales linearly with the number
of tasks, which is a problem if we want to scale to Trinitite's 3000 cores.
Furthermore, the size of the keyspace also increases linearly with the length
of the run.  Extrapolating these results puts an 8 hour run across all 100
Trinitite nodes at 8GB for one cache.  This memory utilization easily eclipses
the 3\% memory usage per node threshold we set earlier, even without factoring
in the usage from other workers.

@@

The bars in Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many
reads (\texttt{get()}) as writes (\texttt{put()}).  Tasks read the same
key for extended periods because the trajectory gets stuck in so-called
superbasins composed of tightly connected sets of states. In this case, many
trajectory segments with the same coordinates are needed before the trajectory
moves on.  Writes only occur for the final state of segments generated by
tasks; their magnitude is smaller than reads because the caches ignore
redundant write requests. The number of read and write requests are highest at
the beginning of the run when tasks generate segments for the same state, which
is computationally cheap (this motivates Section~\S\ref{sec:arch-specific}).

### 2017-11-28 paper/sections/3_methodology.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%  \item background: CephFS implementation
To explore software-defined cache management, we use the data management
language and policy engine presented in~\cite{sevilla:sc15-mantle}. The
prototype in that paper was called Mantle and it was originally built on CephFS
so administrators could control file system metadata load balancing policies.
We now refer to Mantle as a policy engine that supports our data management
language.  The basic premise is that data management policies can be expressed
with a simple API consisting of ``when", ``where", and ``how much". The ``when" policy
controls how aggressive or conservative the decisions are; ``where" controls
how distributed or concentrated the data should be; and ``how much" controls
the amount of data that should be sent. There is also a ``load" policy that
lets administrators specify how to collapse many metrics into a single load
metric ({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).

The succinctness of the API lets users inject multiple, possibly dynamic, policies. In
this work we focus on a single node, so the ``where" policy is not used. When
we move ParSplice to a distributed key-value store back-end, the ``where" policy
will be used determine which key-value pairs should be moved to which node.

@@

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a storage system.  Administrators write policies in Lua from whatever
domain they choose ({\it e.g.}, statistics, machine learning, storage system)
and the policies are embedded into the runtime by Mantle.  We chose Lua for
simplicity, performance, and portability; it is a scripting language with
simple syntax, which allows administrators to focus on the policies themselves;
it was designed as an embeddable language, so it is lightweight and does less
type checking; and it interfaces nicely with C/C++.  When the storage system makes
decisions it executes the administrator-defined policies for when/where/how
much and returns a decision.  To do this, the storage system needs to be modified to
(1) provide an environment of metrics and (2) identify where policies are set.
These modification points are shown by the colored boxes in
Figure~\ref{fig:mantle} and described below.

@@

\subsubsection{Environment of Metrics} storage systems expose \textbf{cluster} metrics
for describing resource utilization and \textbf{time series} metrics for
describing accesses to some data structure over time. Table~\ref{table:metrics}
shows how these metrics are accessed from the policies written by
administrators.

@@

For time series metrics, the storage system passes an array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the accesses.  The
storage system uses a pointer to the time series to facilitate time series with many
values, like accesses to a database or directory in the file system namespace.
This decision limits the time series metrics to only include values from the
{\it current } node, although this is not a limitation of Mantle itself.  For
example, a policy that uses accesses to a directory in a file system as a
metric for load collects that information using:

@@

% where policies are made
We link Mantle directly into the worker cache circled in red in
Figure~\ref{fig:parsplice}. We put the ``when" and ``how much" callbacks
alongside code that checks for memory pressure. It is executed right before the
worker processes incoming and outgoing put/get transactions to the cache. We
only do cache management once every second to avoid maintaining the cache for
every request. We expected to have to increase this polling interval to
accommodate more complex policies but even our most complicated policy in
Section~\S\ref{sec:dom-specific} had a negligible effect on performance when
executed every second (within the standard deviation for multiple runs when
compared against a policy that returns immediately). This may be because the
worker is not overloaded and the bottleneck is somewhere else in the system.
As stated previously, we do not use the ``where" part of Mantle because we
focus on a single node, but this part of the API will be used when we move the
caches and storage nodes to a key-values store back-end that uses key load
balancing and repartitioning.

>>>>

%  \item background: CephFS implementation
To explore software-defined cache management, we use the data management
language and policy engine presented in~\cite{sevilla:sc15-mantle}. The
prototype in that paper, Mantle, was built on CephFS and lets administrators
control file system metadata load balancing policies.  We now refer to Mantle
as a policy engine that supports our data management language.  The basic
premise is that data management policies can be expressed with a simple API
consisting of ``when", ``where", and ``how much". The ``when" policy controls
how aggressive or conservative the decisions are; ``where" controls how
distributed or concentrated the data should be; and ``how much" controls the
amount of data that should be sent. There is also a ``load" policy that lets
administrators specify how to collapse many metrics into a single load metric
({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).

The succinctness of the API lets users inject multiple, possibly dynamic, policies. In
this work we focus on a single node, so the ``where" policy is not used. When
we move ParSplice to a distributed key-value store back-end, the ``where" policy
will be used to determine which key-value pairs should be moved to which node.

@@

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a storage system service.  Administrators write policies in Lua from whatever
domain they choose ({\it e.g.}, statistics, machine learning, storage system)
and the policies are embedded into the runtime by Mantle.  We chose Lua for
simplicity, performance, and portability; it is a scripting language with
simple syntax, which allows administrators to focus on the policies themselves;
it was designed as an embeddable language, so it is lightweight and does less
type checking; and it interfaces nicely with C/C++.  When the storage system makes
decisions it executes the administrator-defined policies for when/where/how
much and returns a decision.  To do this, the storage system needs to be modified to
(1) provide an environment of metrics and (2) identify where policies are set.
These modification points are shown by the colored boxes in
Figure~\ref{fig:mantle} and described below.

@@

\subsubsection{Environment of Metrics} storage systems expose \textbf{cluster} metrics
for describing resource utilizations and \textbf{time series} metrics for
describing accesses to some data structure over time. Table~\ref{table:metrics}
shows how these metrics are accessed from the policies written by
administrators.

@@

For time series metrics, the storage system passes an array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the values. The
storage system uses a pointer to the time series to facilitate time series with many
values, like accesses to a database or directory in the file system namespace.
This decision limits the time series metrics to only include values from the
{\it current } node, although this is not a limitation of Mantle itself.  For
example, a policy that uses accesses to a directory in a file system as a
metric for load collects that information using:

@@

% where policies are made
We link Mantle into all caches in the system and put the ``when" and ``how
much" callbacks alongside code that checks for memory pressure. It is executed
right before the worker processes incoming and outgoing put/get transactions to
the cache. We only do cache management once every second to avoid maintaining
the cache for every request. We expected to have to increase this polling
interval to accommodate more complex policies but even our most complicated
policy in Section~\S\ref{sec:dom-specific} had a negligible effect on
performance when executed every second (within the standard deviation for
multiple runs when compared against a policy that returns immediately). This
may be because the worker is not overloaded and the bottleneck is somewhere
else in the system.  As stated previously, we do not use the ``where" part of
Mantle because we focus on a single node, but this part of the API will be used
when we move the caches and storage nodes to a key-values store back-end that
uses key load balancing and repartitioning.

### 2017-11-28 paper/sections/4_dynamic-policies.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on the worker using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}.  Our evaluation uses the total
``trajectory length" as the goodness metric. This value is the duration of the
overall trajectory produced by ParSplice. At ideal efficiency, the trajectory
length should increase with the square root of the wall-clock time, since the
wall-clock cost of time-stepping the system by one simulation time unit
increases in proportion of the total number of atoms.  The policy should avoid
reducing the trajectory length and be fast enough to run as often as we want to
detect key access patterns.  First we sized the cache according to our system
specific knowledge, {\it i.e.} the hardware and software of the storage
hierarchy such as request rate, unique keys in a sliding window, and bandwidth
capabilities.

% technical details
The results for different cache sizes for a growth rate of \(\Delta_1\) over a
2.5 hour run across 256 workers is shown in
Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
utilization is measured with memory footprint of just the cache (\(y2\) axis).
The middle graph labeled ``Fixed Cache Size" shares the \(y\) axis and shows the
trade-off of using a basic LRU-style cache for different cache sizes, where the
penalty for a cache miss is retrieving the data from the persistent database.
We evict keys (if necessary) at every operation instead of when segments
complete because the cache fills up too quickly otherwise. This is implemented
using a ``when" policy of:

@@

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
\caption{Policy performance/utilization: ``None" is ParSplice unmodified,
``Fixed Sized Cache" evicts keys using LRU , and ``Multi-Policy Cache" switches
to fixed sized cache after absorbing the workload's initial burstiness.  Fixed
sized caches are on the \(x\) axis.\label{fig:methodology-tradeoff}}
\end{figure}

\begin{figure}[t]
        \centering
        \includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
	\caption{Memory utilization for ``No Cache Management" (unlimited cache
growth), ``Multi-Policy" (absorbs initial burstiness of workload) , and
``Dynamic Policy" (sizes cache according to key access patterns).
\label{fig:memory-vs-time}}
\end{figure}%

@@

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
two reasons: (1) segments take longer to generate later in the run and (2) the
memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
length vs.  wall-clock time curves down over time; as the nanoparticle grows it
takes longer to generate segments so by the time we reach 2 hours, over 90\% of
the trajectory is already generated.  For (2), the memory footprint rises until
it reaches the 100K key switch threshold at 0.4GB and then reduces to the final
value after switching policies. The memory usage over time for this policy is
shown by the ``\(\Delta_1\), Multi-Policy" curve in
Figure~\ref{fig:memory-vs-time} but in Figure~\ref{fig:methodology-tradeoff} we
plot the final value.  Despite these caveats, the result is still valid: we
found a multi-policy cache management strategy that absorbs the cost of a high
read throughput on a small keyspace and reduces the memory pressure for a 2.5
hour run. Our experiments show the effectiveness of the policy engine we
integrated into ParSplice, not that we were able to identify the best policy
for all system setups ({\it i.e.} different ParSplice parameters, number of
worker tasks, and job lengths).  To solve that problem, we need a way to
identify what thresholds we should use for different configurations.

@@

\begin{figure*}[t!]
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
\caption{Key activity for a 4 hour run shows groups of accesses to the same
subset of keys. Detecting these access patterns leads to a more accurate cache
management strategy, which is discussed in
Section~\S\ref{sec:regime-detection}.\label{fig:keyspace-zoomed}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
	\caption{Performance/Utilization for dynamically sized cache (DSCache)
policy. With negligible performance degradation, DSCache adjusts to different
\(\Delta\)s over 2\(\times\) memory in the best case.
\label{fig:dscache-vs-none}}
    \end{subfigure}%
    \caption{Different cache management policies tested over the Mantle policy engine.}
\end{figure*}

@@

Feeding application-specific knowledge about the ParSplice application into a policy
leads to a more accurate cache management strategy.  The goal of the following
sections is not to find an optimal solution, as this can be done with parameter
sweeps for thresholds; rather, we try to find techniques that work for a range
of inputs and system setups.

@@

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that the following
techniques proliferate more parameters that need to be tuned per
hardware/software configuration. Furthermore, we find that many of the metrics
do not signal a new set of key accesses.  Below, we indicate with quotes which
parameters we need to add for each technique with and the value we find to work
best, via tuning and parameter sweeps, for one set of initial conditions.

@@

  \item Anomaly Detection: size of the image is too big and bottom edges are
not thick enough.

@@

After trying these techniques we found that the basic algorithm in
Figure~\ref{src:dyn-cache} works best. The algorithm detects groups of key
accesses, which we call ``fans", by iterating backwards through the key access
trace, finding the lowest key ID, and comparing against the lowest key ID we
have seen so far (Line 7). We also maintain the top and bottom of each group of
key accesses (Line 13) so we can tell the ``how much" policy the number of keys
to evict (Line 19).  The algorithm iterates backwards over the key access trace
because a change in the minimum value signals a new group of key accesses. No
signal exists iterating left to right, as the maximum value always increases
and the minimum values at the bottom of each group of key accesses are sparse.
For example, the maximum distance between values along the bottom edge of the
zoomed group of key accesses in Figure~\ref{fig:keyspace-zoomed} is 125
seconds, while the maximum distance between minimum values for the group of key
accesses before is 0 seconds. As a result of this sparseness, iterating left to
right requires a ``window size" parameter to determine when we think a minimum
value will not show up again.  The algorithm is O(\(n\)), where \(n\) is the
number events, but the benefit is that the approach avoids adding new
thresholds for key access pattern detection ({\it e.g.}, space between key
accesses, space between key IDs, and window size of consecutive key accesses).

The performance and memory utilization is shown by the ``DSCache" bars in
Figure~\ref{fig:dscache-vs-none}. Without sacrificing performance (trajectory
length), the dynamically sized cache policy uses between 32\%-66\% less memory
than the default ParSplice configuration (no cache management) for the 3
initial conditions we test. The memory usage is shown by the ``Dynamic Policy"
curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
access patterns in Figure~\ref{fig:keyspace-zoomed}\footnote{The memory usage
is not {\it exactly} the same because these are two different runs;
Figure~\ref{fig:keyspace-zoomed} has key activity tracing turned on, which
reduces performance.}.  We also show a \(\Delta_2\) growth rate to demonstrate
the dynamic policy's ability to adjust to a different set of initial
conditions.

\begin{figure}[h]
\footnotesize
\begin{minted}[xleftmargin=3em,linenos]{lua}
d = timeseries()
ts, id = d:get(d:size())
fan  = {start=nil, finish=ts, top=0, bot=id}
fans = {}
for i=d:size(),1,-1 do     -- iterate backwards
  ts, id = d:get(i)
  if id < fan['bot'] then  -- found a new fan!
    fan['start']  = ts
    fans[#fans+1] = fan 
    fan = {start=nil, finish=ts, top=0, bot=id}
  end

>>>>

Using the Mantle policy engine, we test a variety of cache management 
algorithms on the worker using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}.  Our evaluation uses the total
``trajectory length" as the goodness metric. This value is the duration of the
overall trajectory produced by ParSplice. At ideal efficiency, the trajectory
length should increase with the square root of the wall-clock time, since the
wall-clock cost of time-stepping the system by one simulation time unit
increases in proportion of the total number of atoms.  The policy should avoid
reducing the trajectory length and be fast enough to run as often as we want to
detect key access patterns.  First we size the cache according to our system
specific knowledge, {\it i.e.} the hardware and software of the storage
hierarchy.

% technical details
The results for different cache sizes for a growth rate of \(\Delta_1\) over a
2.5 hour run across 256 workers is shown in
Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
utilization is measured with memory footprint of just the cache (\(y2\) axis).
The middle graph labeled ``Fixed Cache Size" shares the \(y\) axes and shows the
trade-off of using a basic LRU-style cache of different sizes, where the
penalty for a cache miss is retrieving the data from the persistent database.
We evict keys (if necessary) at every operation instead of when segments
complete because the cache fills up too quickly otherwise. This is implemented
using a ``when" policy of:

@@

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
\caption{Policy performance/utilization shows the trade-offs of different sized caches
(\(x\) axis).  ``None" is ParSplice unmodified, ``Fixed Sized Cache" evicts
keys using LRU, and ``Multi-Policy Cache" switches to fixed sized cache after
absorbing the workload's initial burstiness.  \label{fig:methodology-tradeoff}}
\end{figure}

\begin{figure}[t]
        \centering
        \includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
	\caption{Memory utilization for ``No Cache Management" (unlimited cache
growth), ``Multi-Policy" (absorbs initial burstiness of workload), and
``Dynamic Policy" (sizes cache according to key access patterns). The ``Dynamic
Policy" saves the most memory without sacrificing performance.
\label{fig:memory-vs-time}}
\end{figure}%

@@

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
two reasons: (1) segments take longer to generate later in the run and (2) the
memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
length vs.  wall-clock time curves down over time; as the nanoparticle grows it
takes longer to generate segments so by the time we reach 2 hours, over 90\% of
the trajectory is already generated.  For (2), the memory footprint rises until
it reaches the 100K key switch threshold at 0.4GB and then reduces to the final
value after switching policies. The memory usage over time for this policy is
shown by the ``\(\Delta_1\), Multi-Policy" curve in
Figure~\ref{fig:memory-vs-time} but in Figure~\ref{fig:methodology-tradeoff} we
plot the final value.  Despite these caveats, the result is still valid: we
found a multi-policy cache management strategy that absorbs the cost of a high
read throughput on a small keyspace and reduces the memory pressure for a 2.5
hour run.  To improve the policy even more, we need a way to identify what
thresholds to use for different system setups ({\it e.g.}, different ParSplice
parameters, number of worker tasks, and job lengths).

@@

\begin{figure*}[t!]
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\

\caption{Key activity for a 4 hour run shows groups of accesses to the same
subset of keys. Detecting these access patterns leads to a more accurate cache
management strategy, which is discussed in Section~\S\ref{sec:regime-detection}
and the results are in Figure~\ref{fig:dscache-vs-none}.
\label{fig:keyspace-zoomed}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\

	\caption{Performance/Utilization for dynamically sized cache (DSCache)
policy. With negligible performance degradation, DSCache adjusts to different
\(\Delta\)'s and saves 3\(\times\) as much memory in the best case.
\label{fig:dscache-vs-none}}

    \end{subfigure}%
    \caption{Different cache management policies tested over the Mantle policy engine.}
\end{figure*}

\begin{figure}[tb]
\footnotesize
\begin{minted}[xleftmargin=3em,linenos]{lua}
d = timeseries()
ts, id = d:get(d:size())
fan  = {start=nil, finish=ts, top=0, bot=id}
fans = {}
for i=d:size(),1,-1 do     -- iterate backwards
  ts, id = d:get(i)
  if id < fan['bot'] then  -- found a new fan!
    fan['start']  = ts
    fans[#fans+1] = fan 
    fan = {start=nil, finish=ts, top=0, bot=id}
  end

@@

Feeding application-specific knowledge about ParSplice into a policy
leads to a more accurate cache management strategy.  The goal of the following
section is not to find an optimal solution, as this can be done with parameter
sweeps for thresholds; rather, we try to find techniques that work for a range
of inputs and system setups.

@@

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that the following
techniques proliferate more parameters that need to be tuned per
hardware/software configuration. Furthermore, many of the metrics do not signal
a new set of key accesses.  Below, we indicate with quotes which parameters we
need to add for each technique and the value we find to work best, via
tuning and parameter sweeps, for one set of initial conditions.

@@

  \item Edge Detection: size of the image is too big and bottom edges are
not thick enough.

@@

After trying these techniques we found that the basic algorithm in
Figure~\ref{src:dyn-cache} works best. The algorithm detects groups of key
accesses, which we call ``fans", by iterating backwards through the key access
trace, finding the lowest key ID, and comparing against the lowest key ID we
have seen so far (Line 7). We also maintain the top and bottom of each group of
key accesses (Line 13) so we can tell the ``how much" policy the number of keys
to evict (Line 19).  The algorithm is O(\(n\)), where \(n\) is the number
events, but the benefit is that the approach avoids adding new thresholds for
key access pattern detection ({\it e.g.}, space between key accesses, space
between key IDs, and window size of consecutive key accesses).

The algorithm iterates backwards over the key access trace because a change in
the minimum value signals a new group of key accesses. No signal exists
iterating left to right, as the maximum value always increases and the minimum
values at the bottom of each group of key accesses are sparse.  For example,
the maximum distance between values along the bottom edge of the zoomed group
of key accesses in Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the
maximum distance between minimum values for the group of key accesses before is
0 seconds. As a result of this sparseness, iterating left to right requires a
``window size" parameter to determine when we think a minimum value will not
show up again.

The performance and memory utilization is shown by the ``DSCache" bars in
Figure~\ref{fig:dscache-vs-none}. Without sacrificing performance (trajectory
length), the dynamically sized cache policy uses between 32\%-66\% less memory
than the default ParSplice configuration (no cache management) for the 3
initial conditions we test. The memory usage over time is shown by the ``Dynamic Policy"
curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
access patterns in Figure~\ref{fig:keyspace-zoomed}\footnote{The memory usage
is not {\it exactly} the same because these are two different runs;
Figure~\ref{fig:keyspace-zoomed} has key activity tracing turned on, which
reduces performance.}.  We also show a \(\Delta_2\) growth rate to demonstrate
the dynamic policy's ability to adjust to a different set of initial
conditions.
